Rapid technological progress in white light-emitting diodes \\ and its sources in innovation and technology spillovers

### 2023-01-21 2_SSL_EES/article/article_main.tex
added content to supplementary information

\subsection{Performance Metrics Calculations}

@@

Some text here

@@

\begin{figure*}[h!]
 \centering
 \includegraphics[width=17.5cm]{2_SSL_EES/article/figures/subefficiencies_progress.pdf}
 \caption{Historical improvements in light-emitting diode technology. Panels A1-A2 show they key metrics used to compute forward voltage efficiency $\eta_{Vf}$ and droop efficiency $\eta_{droop}$. Panels B1-B2 show the sub-efficiencies internal quantum efficiency $\eta_{IQ}$ and light extraction efficiency $\eta_{LE}$. \newline
 \textbf{Panel A1:} Device forward voltage at a test current of I=350mA. The physical limit for a blue light wavelength of 450nm without electric pumping is shown for reference. Data points for devices released in 2020 by various manufacturers are shown in an inset plot.
 \textbf{Panel A2:} Efficiency droop at the test current of I=350mA.
 \textbf{Panel B1:} Internal quantum efficiency, for different chip architectures, by type of measurement used. Measurement methods: PL - Photoluminescence\cite{Shim_2018}, EL - Electroluminescence\cite{Getty_2009};  ABC Model\cite{Karpov_2014}, Ideality Factor\cite{Masui_2010}. “Not reported” denotes data points where the IQE measurement technique was not reported. Whiskers indicate reported ranges, where applicable.
 \textbf{Panel B2:} Light extraction efficiency for different chip architectures, by type of specific technology used to improve light-extraction efficiency. Abbreviations: TFFC - Thin-Film Flip-Chip; FC - Flip-Chip; CSP - Chip-Scale Package; ITO - Indium Tin Oxide; ZnO - Zinc Oxide; PSS - Patterned Sapphire Substrate; electr. - electrode; text. - textured.
 Source: own synthesis of data from the full list of sources provided in the Supplementary Information, section XX. Data on average device performance adapted from U.S. Department of Energy (DOE) Reports\cite{doe_ssl_multiyear_2007}\cite{doe_ssl_multiyear_2008}\cite{doe_ssl_multiyear_2013}\cite{doe_ssl_rnd_2016}\cite{doe_ssl_rnd_2018} and an Osram company report\cite{beale_leds_2015}.}
 \label{fgr:waterfall}
\end{figure*}

@@

Historical improvements in consumer experience metrics for phosphor-converted warm white LEDs are shown in Figure \ref{fgr:consumer_experience}. A high colour rendering index (CRI) of an LED in combination with a specific range of possible colour temperatures is generally desirable in general illumination applications. The values of these metrics are determined by LED device spectra (provided and discussed in SI section X), which, in turn, depend on the means of light down conversion in the device. Due to that, we were able to establish the links between all major improvements in LED consumer experience metrics identified in this study and individual LED inventions associated either with phosphors or quantum dots used for light down conversion. The list of these inventions is provided in Table 3.

@@

\begin{table*}[h]
    \small
    \centering
    \begin{tabularx}{\textwidth}{|l|l|l|X|X|l|}
    \hline
        \textbf{Year} & \textbf{Desig.} & \textbf{Chemical formula} & \textbf{Description} & \textbf{Significance} & \textbf{SI Sec.} \\ \hline
        1996 & YAG & $Y_3 Al_5 O_{12}:Ce$ & Yttrium aluminium garnet (YAG) phosphor activated with cerium & First LED phosphor, enabled white LEDs & SI 3.1 \\ \hline
        1996 & YGAG & $(Y_{1-x} Gd_x)_3 Al_5 O_{12}:Ce$ & Gadolinium-doped YAG phosphor & First red-shifted phosphor, enabled warm white LEDs & SI 3.1 \\ \hline
        2002 & 258 & $(Ba,Sr)_2 Si_5 N_8:Eu^{2+}$ & Europium-doped nitridosilicate phosphor & First red LED phosphor & SI 3.2 \\ \hline
        2003 & QD & N/A & Quantum dot-based phosphor & First use of QD for LED light down conversion & SI 3.4 \\ \hline
        2005 & PFS & $K_2 SiF_6: Mn^{4+}$ & Manganese-activated potassium fluorosilicate (PFS) phosphor & First ultra-narrow-band red LED phosphor & SI 3.3 \\ \hline
        2013 & SLA & $Sr[Li Al]_3 N_4 ]:Eu^{2+}$ & Europium-doped cuboidal nitridolithoaluminate phosphor & Improved narrow-band red phosphor & SI 3.2 \\ \hline
        2016 & SALON & $Sr[Li_2 Al_2 O_2 N_2]:Eu^{2+}$ & Europium-doped oxonitride phosphor & High-performance ultra-narrow-band red phosphor & SI 3.2 \\ \hline
    \end{tabularx}
    \caption{LED light down conversion inventions related to improvements in consumer experience metrics identified in this study. Note: Year column represents the earliest reported application of invention in white LEDs. These years are different from the years in Figure \ref{fgr:consumer_experience} that correspond to the earliest publication of spectral data for representative LED products  that relied on those inventions. Section in SI refers to a section in the Supplementary Information that contains references and a detailed description of the history of corresponding invention. Abbreviations: Desig. - Designation, Sec. - Section.}
    \label{tab:phosphors}
\end{table*}

\begin{figure*}[h]
 \centering
 \includegraphics[width=\textwidth]{2_SSL_EES/article/figures/breakthroughs_consumer-experience.pdf}
 \caption{Historical improvements in consumer experience metrics of phosphor-converted white LEDs. Circle dots show data for the earliest identified representative white LED products with published spectral data that used phosphors listed in Table 3, each indicated by the phosphor label and publication year. Corresponding spectral data used to calculate colour temperature values is provided in Figure SI1 in SI section 2. Three data points representing compact fluorescent lamps (CFLs)\cite{cie_reference}, shown as black triangles, are provided for comparison. Horizontal lines represent typical colour temperatures of “traditional” light sources, shown for reference. Desirable range of colour temperatures for home illumination, indicated by a vertical red arrow, lies between two horizontal orange lines representing typical incandescent light and warm daylight colour temperatures. Horizontal red arrow indicates desirable higher values of colour rendering index (CRI). LED products based on the following phosphor innovations from the following LED manufacturers are represented : YAG - Nichia, 1998 \cite{bando1998development}; YGAG - Lumileds, 2002 \cite{Mueller2002}; 258 - Lumileds, 2005 \cite{MuellerMach2005}; SLA - Lumileds, 2014 \cite{Pust2014}; PFS - GE, 2015 \cite{Murphy2015}; QD - Lumileds, 2017 \cite{lumileds2016qd}, Osram, 2019 \cite{osram2019qd}; SALON - Osram, 2019 \cite{Hoerder2019}.}
 \label{fgr:consumer_experience}
\end{figure*}

@@

\subsection{Technology Spillovers}

@@

\section*{Conflicts of interest}
There are no conflicts to declare.

>>>>

\subsection{Performance Metrics Calculations}
\label{subsec:metrics}

@@

@@

The historical improvements in four sub-efficiency channels—forward voltage efficiency, internal quantum efficiency, droop, and light extraction efficiency—identified in our study are shown in Figure 4. Using data on these four sub-efficiencies, as well as on improvements in spectral efficiency and light conversion efficiencies calculated from spectral data for LED phosphors provided in SI section 2 and discussed in the context of consumer experience improvements in section 5.2, we then calculate the overall LED efficiency in 2003 and 2010, shown in Figure XX. as a waterfall diagram across all sub-efficiencies, and in 2016 and 2020, similarly shown in Figure . The overall white LED lamp efficiency improved from $\eta_L(2003)=5.8\% \rightarrow \eta_L (2010) = 12.6\% \rightarrow \eta_L = 32.5\% \rightarrow \eta_L(2020) = 38.7\%$, as shown in Figure \ref{fgr:waterfall}. No single loss channel dominates in terms of its contribution to the overall efficiency, in line with previous observations\cite{tsao2010solid}. We note, however, that the loss channels with a fixed physical limit, e.g., the Stokes loss that contributes to the conversion efficiency by phosphors, became more dominant in 2016 and 2020 compared to 2003 and 2010.

Figure \ref{fgr:breakthroughs_efficiency} shows the overall magnitude of contributions of identified LED innovations and technology spillovers to improvements in LED efficiency over time across different sub-efficiencies. The list of technology spillovers identified in our study is provided and discussed below in Section \ref{subsec:spillovers}. Through the index decomposition analysis described in section \ref{subsec:metrics}, we find that in the overall LED efficiency increase of 32.9\% from 5.8\% to 38.7\% between 2003 and 2020, at least 2.8 of these percentage points can be attributed specifically to technology spillovers identified in this study, corresponding to 8.5\% of the total LED efficiency improvements between 2003 and 2020.

In Figure \ref{fgr:breakthroughs_efficiency} we also compare, for the first time, efficiency improvements across sub-efficiencies over time, contrasting them with the physical limits of corresponding loss channels. Notably, sub-efficiencies for the most recent devices considered in our study are only $\sim10\%$ below the physical limit. The exception is spectral efficiency, which at $\sim20\%$ below the physical limit shows larger potential for further improvements.

\begin{figure*}[h!]
 \centering
 \includegraphics[width=17.5cm]{2_SSL_EES/article/figures/subefficiencies_progress.pdf}
 \caption{Historical improvements in light-emitting diode technology. Panels A1-A2 show they key metrics used to compute forward voltage efficiency $\eta_{Vf}$ and droop efficiency $\eta_{droop}$. Panels B1-B2 show the sub-efficiencies internal quantum efficiency $\eta_{IQ}$ and light extraction efficiency $\eta_{LE}$. \newline
 \textbf{Panel A1:} Device forward voltage at a test current of I=350mA. The physical limit for a blue light wavelength of 450nm without electric pumping is shown for reference. Data points for devices released in 2020 by various manufacturers are shown in an inset plot.
 \textbf{Panel A2:} Efficiency droop at the test current of I=350mA.
 \textbf{Panel B1:} Internal quantum efficiency, for different chip architectures, by type of measurement used. Measurement methods: PL - Photoluminescence\cite{Shim_2018}, EL - Electroluminescence\cite{Getty_2009};  ABC Model\cite{Karpov_2014}, Ideality Factor\cite{Masui_2010}. “Not reported” denotes data points where the IQE measurement technique was not reported. Whiskers indicate reported ranges, where applicable.
 \textbf{Panel B2:} Light extraction efficiency for different chip architectures, by type of specific technology used to improve light-extraction efficiency. Abbreviations: TFFC - Thin-Film Flip-Chip; FC - Flip-Chip; CSP - Chip-Scale Package; ITO - Indium Tin Oxide; ZnO - Zinc Oxide; PSS - Patterned Sapphire Substrate; electr. - electrode; text. - textured.
 Source: own synthesis of data from the full list of sources provided in the Supplementary Information, section XX. Data on average device performance adapted from U.S. Department of Energy (DOE) Reports\cite{doe_ssl_multiyear_2007}\cite{doe_ssl_multiyear_2008}\cite{doe_ssl_multiyear_2013}\cite{doe_ssl_rnd_2016}\cite{doe_ssl_rnd_2018} and an Osram company report\cite{beale_leds_2015}.}
 \label{fgr:subeff}
\end{figure*}

@@

Historical improvements in consumer experience metrics for phosphor-converted warm white LEDs are shown in Figure \ref{fgr:consumer_experience}. A high colour rendering index (CRI) of an LED in combination with a specific range of possible colour temperatures is generally desirable in general illumination applications. The values of these metrics are determined by LED device spectra (provided and discussed in SI section X), which, in turn, depend on the means of light down conversion in the device. Due to that, we were able to establish the links between all major improvements in LED consumer experience metrics identified in this study and individual LED inventions associated either with phosphors or quantum dots used for light down conversion. The list of these inventions is provided in Table \ref{tab:spillovers}.

@@

\begin{table*}[h!]
    \small
    \centering
    \begin{tabularx}{\textwidth}{|l|l|l|X|X|l|}
    \hline
        \textbf{Year} & \textbf{Desig.} & \textbf{Chemical formula} & \textbf{Description} & \textbf{Significance} & \textbf{SI Sec.} \\ \hline
        1996 & YAG & $Y_3 Al_5 O_{12}:Ce$ & Yttrium aluminium garnet (YAG) phosphor activated with cerium & First LED phosphor, enabled white LEDs & SI 3.1 \\ \hline
        1996 & YGAG & $(Y_{1-x} Gd_x)_3 Al_5 O_{12}:Ce$ & Gadolinium-doped YAG phosphor & First red-shifted phosphor, enabled warm white LEDs & SI 3.1 \\ \hline
        2002 & 258 & $(Ba,Sr)_2 Si_5 N_8:Eu^{2+}$ & Europium-doped nitridosilicate phosphor & First red LED phosphor & SI 3.2 \\ \hline
        2003 & QD & N/A & Quantum dot-based phosphor & First use of QD for LED light down conversion & SI 3.4 \\ \hline
        2005 & PFS & $K_2 SiF_6: Mn^{4+}$ & Manganese-activated potassium fluorosilicate (PFS) phosphor & First ultra-narrow-band red LED phosphor & SI 3.3 \\ \hline
        2013 & SLA & $Sr[Li Al]_3 N_4 ]:Eu^{2+}$ & Europium-doped cuboidal nitridolithoaluminate phosphor & Improved narrow-band red phosphor & SI 3.2 \\ \hline
        2016 & SALON & $Sr[Li_2 Al_2 O_2 N_2]:Eu^{2+}$ & Europium-doped oxonitride phosphor & High-performance ultra-narrow-band red phosphor & SI 3.2 \\ \hline
    \end{tabularx}
    \caption{LED light down conversion inventions related to improvements in consumer experience metrics identified in this study. Note: Year column represents the earliest reported application of invention in white LEDs. These years are different from the years in Figure \ref{fgr:consumer_experience} that correspond to the earliest publication of spectral data for representative LED products  that relied on those inventions. Section in SI refers to a section in the Supplementary Information that contains references and a detailed description of the history of corresponding invention. Abbreviations: Desig. - Designation, Sec. - Section.}
    \label{tab:phosphors}
\end{table*}

\begin{figure*}[h!]
 \centering
 \includegraphics[width=\textwidth]{2_SSL_EES/article/figures/breakthroughs_consumer-experience.pdf}
 \caption{Historical improvements in consumer experience metrics of phosphor-converted white LEDs. Circle dots show data for the earliest identified representative white LED products with published spectral data that used phosphors listed in Table 3, each indicated by the phosphor label and publication year. Corresponding spectral data used to calculate colour temperature values is provided in Figure SI1 in SI section 2. Three data points representing compact fluorescent lamps (CFLs)\cite{cie_reference}, shown as black triangles, are provided for comparison. Horizontal lines represent typical colour temperatures of “traditional” light sources, shown for reference. Desirable range of colour temperatures for home illumination, indicated by a vertical red arrow, lies between two horizontal orange lines representing typical incandescent light and warm daylight colour temperatures. Horizontal red arrow indicates desirable higher values of colour rendering index (CRI). LED products based on the following phosphor innovations from the following LED manufacturers are represented : YAG - Nichia, 1998 \cite{bando1998development}; YGAG - Lumileds, 2002 \cite{Mueller2002}; 258 - Lumileds, 2005 \cite{MuellerMach2005}; SLA - Lumileds, 2014 \cite{Pust2014}; PFS - GE, 2015 \cite{Murphy2015}; QD - Lumileds, 2017 \cite{lumileds2016qd}, Osram, 2019 \cite{osram2019qd}; SALON - Osram, 2019 \cite{Hoerder2019}.}
 \label{fgr:consumer_experience}
\end{figure*}

@@

\subsection{Technology Spillovers}
\label{subsec:spillovers}

@@

\section*{Conflicts of interest}

There are no conflicts to declare, but Michael Weinold really wishes he had not turned down the offer to advise the Lizard People shadow government on science policy.

### 2023-01-21 2_SSL_EES/article/article_supplementary.tex
added content to supplementary information

@@

\begin{figure}[h]
	\centering
    \includegraphics[width=\textwidth]{2_SSL_EES/article/figures/costmodel_sensitivity.pdf}
	\caption{Panels a-b: Sensitivity analysis of selected cost model parameters. Note the decreasing sensitivity to cost model parameters as the number of die per wafer increases.}
	\label{fig:sensitivity}
\end{figure}

@@

>>>>

\subsection{Examples of Input Data Considered}

\subsubsection{Sapphire Wafers}

Sapphire wafers form the substrate on which all other layers of the light-emitting diode wafer are grown. Being transparent to radiation in the visible spectrum, it is not removed after growth in the Classical architecture or the Thin-Film Flip-Chip architecture. In the Vertical Thin-Film architecture, it is removed by means of a laser-lift-off process. Wafers can be either unpatterned or patterned, where the latter has become commonplace in 2020 due to the beneficial properties that microstructures on the surface have on layer growth \cite{wuu2009defect} and light-extraction efficiency \cite{lee2006enhancing}. The price of sapphire substrates has decreased significantly since the year 2000, as shown in figure \ref{fig:sapphire_prices}. This can be attributed not only to the lighting industry, but more importantly to increased demand from electronics manufacturing \cite{yole2015sapphire}, where sapphire glass is used to protect screens and sensor interfaces from scratches \cite{khattak2016world}. Wafer sizes used in manufacturing have also increased, due to the favourable economics of large wafer processing. The market is dominated by U.S. based \textit{Rubicon Technology} and Russian based \textit{Monocrystal}. The prevalence of wafer diameters used is shown in figure \ref{fig:wafer_size}.

\begin{figure}[h!]
    \includegraphics[width=14.5cm]{./figures/wafer_size.pdf}
    \includegraphics[width=15cm]{./figures/sapphire_prices.pdf}
	\caption{Top: Historical data for sapphire substrate prices of different surface properties and diameters. Shown is data for polished surface (NPS) and patterned substrates (PSS). \textit{Monocrystal} denotes the Russian manufacturer of the same name. Dashed lines are projections from the previous year. Sources: \cite{monocrystal2020private}\cite{yole2011sapphire}\cite{yole2015sapphire}. Bottom: Prevalence of sapphire wafer size used in the manufacturing of light-emitting diodes. Hatched bars are projections given by the sources. Sources: \cite{veeco2013}\cite{Scholand2012}\cite{yole2015sapphire}}
	\label{fig:wafer_size}
\end{figure}

@@

\begin{figure}[h]
	\centering
    \includegraphics[width=\textwidth]{2_SSL_EES/article/figures/costmodel_sensitivity.pdf}
	\caption{Sensitivity analysis of selected cost model parameters. Note the decreasing sensitivity to cost model parameters as the number of die per wafer increases. Compare table \ref{tab:sensitivity}. Abbreviations: resp. - respective}
	\label{fig:sensitivity}
\end{figure}

@@

Figure \ref{fig:costmodel_calibration} shows a comparison of the relative manufacturing cost by manufacturing step categories between the cost model and data released by the United States Department of Energy (DoE). Note that any comparison between the figures must consider: Our model assumes state-of-the-art equipment, while the DoE industry round table reports data from factories which use older equipment, as well as newer equipment. Our model presently considers only low-power and mid-power packages, while the DoE reports consider high-power packages also. Our model assumes a virtual North American manufacturing location, while the majority of manufacturers produce in Asia. We note that the share of the substrate price in the DoE data is much larger than in our model. This can in part be explained by figure \ref{fig:sapphire_prices}, which already shows how projections overestimated the actual price for 2015-2020. We note also that the share of the epitaxy step is larger in the DoE model. This can in part be explained by our model relying on state-of-the-art equipment, while industry might not run low-power and mid-power chip production on these, more expensive, reactors. In conclusion, our model fits industry forecasts due to the way it is set up. The packaging part of the manufacturing process is increasing in importance, which has been confirmed by researchers and industry reports on WLP\cite{Lee2011WPL}\cite{Xie2013}\cite{ledsmag2017WLP}.
